What is the opposite of a song?

### 2014-01-25 trunk/midinverse/paper.tex
Add ability to permute notes within an octave

\usepackage{url}
\usepackage{code}
% \usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{graphicx}

@@

\paragraph{MIDI.} It stands for Musical Instrument Digital Interface
and you can tell from the need to clarify that it is ``digital'' that
it is as old as old dead dry dust and just barely adequate for modern
applications, and so has never been successfully superseded. It
permits 16 {\it channels}, each more-or-less a separate instrument,
with 128 {\it notes} per channel. Each of the 128 MIDI notes has a
fixed meaning, from 0 = C1 \comment{XXX?}, which is 5 octaves below
middle C (about 8 Hz\footnote{A note like ``middle C'' has a fixed
  semantics in some wishy-washy music sense, but the actual frequency
  of the tone is determined by mapping that meaning into one of
  several conventional scales. For example, in ``concert pitch'',
  middle A is fixed to 220 Hz (so middle C is XXX Hz), but there is
  also ``computer science pitch'', where middle C is 256 Hz. The
  techniques here are mostly insensitive to the absolute pitch values,
  but we use computer science pitch because what the hey. }) to 127 =
G11 (about 12,274 Hz). A track consists of a time-ordered series of
events, with the main events being ``turn note $n$ on with at volume
$v$'' and ``turn note $n$ off''\footnote{I still haven't decided how I
  feel about this, but the note-off event also includes a velocity,
  which I guess means something like ``how hard you yank your finger
  off the piano key''. It is usually ignored and in fact the
  convention actually seems to encode note-off as ``note-on with
  volume 0''. In this paper I just use note-off to mean whatever
  concrete event results in the note being inaudible.}; volumes also
range 0--127, allowing for every possible nuance of expression. Notes
can be turned on and off arbitrarily, but we restrict ourselves to the
sensible subset where a note-on event always precedes exactly one
note-off event for the same note on the same track, with no nesting or
overlaps or other shenanigans. For such MIDIs, there exists a simple
isomorphism between the event-based representation and a ``bitmap''
representation where for each MIDI tick, each note in each track has a
volume level 0--127, where 0 is predominant and means ``off''. Of course
there are a jillion other oddities to MIDI, but this will suffice as
a definition of a ``song'' for now.

@@

There are many formal and informal ideas of ``opposite''. Here I will
use a very loose interpretation of opposite so that we can explore
many different reasonable ideas. We define not the opposite of a
specific thing, but a method for getting the opposite of all of the
things of some type. So an opposite of some thing of type $A$ will be a
function $f$ of type $A \rightarrow A$ where
%
\begin{itemize}
\item $\exists x{:}A. f(x) \neq x$
\item $\forall x{:}A. f(f(x)) = x$
\end{itemize}

@@

The second condition is that f is its own inverse, which expresses
the idea that items ``pair up'' into their opposites. So if the opposite
of day is night, then we expect the opposite of night to be day.

@@

(XXX about what this idea is in mathematics--is it an involution?)

>>>>

\usepackage{url}
% \usepackage{code}
% \usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

@@

\paragraph{MIDI.} It stands for Musical Instrument Digital Interface
and you can tell from the need to clarify that it is ``digital'' that
it is as old as old dead dry dust and just barely adequate for modern
applications, and so has never been successfully superseded. It
permits 16 {\it channels}, each more-or-less a separate
instrument,\!\footnote{It's really more like 16 members of an
  orchestra, each of whom can only play one instrument at a time, but
  we'll make the additional simplification in this paper that each
  channel corresponds to a single instrument.} with 128 {\it notes}
per channel. Each of the 128 MIDI notes has a fixed meaning, from 0 =
C$_{-1}$, which is 5 octaves below middle C (about 8~Hz\footnote{A
  note like ``middle C'' has a fixed semantics in some wishy-washy
  music sense, but the actual frequency of the tone is determined by
  mapping that meaning into one of several conventional scales. For
  example, in ``concert pitch'', middle A is fixed to 220 Hz (so
  middle C is 261.63~Hz), but there is also ``computer science
  pitch'', where middle C is 256~Hz. The techniques here are mostly
  insensitive to the absolute pitch values, but we use computer
  science pitch because what the hey. }) to 127 = G9 (about
12,274~Hz). A track consists of a time-ordered series of events, with
the main events being ``turn note $n$ on with at volume $v$'' and
``turn note $n$ off'';\footnote{I still haven't decided how I feel
  about this, but the note-off event also includes a velocity, which I
  guess means something like ``how hard you yank your finger off the
  piano key''. It is usually ignored and in fact the convention
  actually seems to encode note-off as ``note-on with volume 0''. In
  this paper I just use note-off to mean whatever concrete event
  results in the note being inaudible.} volumes also range 0--127,
allowing for every possible nuance of expression. Notes can be turned
on and off arbitrarily, but we restrict ourselves to the sensible
subset where a note-on event always precedes exactly one note-off
event for the same note on the same track, with no nesting or overlaps
or other shenanigans. For such MIDIs, there exists a simple
isomorphism between the event-based representation and a ``bitmap''
representation where for each MIDI tick, each note in each track has a
volume level 0--127, where 0 is predominant and means ``off''. Of
course there are a jillion other oddities to MIDI, but this will
suffice as a definition of a ``song'' for now.

@@

There are many formal and informal ideas of ``opposite''. Here I will
use a very loose interpretation of opposite so that we can explore
many different reasonable ideas. We define not the opposite of a
specific thing, but a method for getting the opposite of all of the
things of some type. So an opposite of some thing of type $A$ will be a
total function $f$ of type $A \rightarrow A$ where
%
\begin{itemize}
\item $\exists x{:}A. f(x) \neq x$
\item $\forall x{:}A. f(f(x)) = x$
\end{itemize}

@@

The second condition is that f is its own inverse, which expresses
the idea that items ``pair up'' into their opposites. So if the opposite
of day is night, then we expect the opposite of night to be day. This
is called an ``involution'' in mathematics.

@@
